2024 US Presidential Election Model

### 2024-10-20 paper/paper.qmd
committing changes to paper.qmd and rendered file

```{r}
#| label: fig-averagetable
#| fig-cap: Summary Table of Polling Averages by State
#| echo: false

@@

average_table
```

@@

```{r}
#| label: fig-barplot
#| fig-cap: Average Poll Percentage by State
#| echo: false

@@

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

@@

### Predicted Election Winner By State
```{r}
#| label: fig-statemap
#| fig-cap: Predicted Election Winner by State
#| echo: false
#| warning: false
#| message: false

#### Workspace setup ####
library(tidyverse)
library(usmap)
library(ggplot2)

# Remove any states with missing polling data for Trump or Harris
state_average <- state_average %>%
  filter(!is.na(Trump_pct) & !is.na(Harris_pct))

# Proceed with the prediction for all states
state_average <- state_average %>%
  mutate(predicted_prob = predict(logistic_model, newdata = state_average, type = "response"))

@@

#### Map Visualization ####
plot_usmap(data = state_average, values = "predicted_winner", regions = "states") +
  scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue"), name = "Predicted Winner") +
  labs(title = "Predicted Winner of 2024 US Presidential Election by State") +
  theme(legend.position = "right")
```

@@

knitr::kable(summary_table, caption = "Summary of Predicted 2024 US Presidential Election Results by State")

@@

```{r}
``{r}
#| echo: false
#| eval: true
#| label: tbl-modelsummary
#| tbl-cap: "Explanatory models of election outcome based on polling percentages for Trump and Harris"
#| warning: false

#### Model Summary Table ####
# Create and display a model summary table
modelsummary::modelsummary(
  list("First Model" = logistic_model),
  statistic = c("std.error", "p.value"),  # Use standard error and p-value
  fmt = 2,  # Format numbers to 2 decimal places
  output = "markdown"  # Output format for Quarto or RMarkdown
)

>>>>

```{r}
#| label: fig-averagetable
#| fig-cap: Summary Table of Polling Averages by State
#| echo: false
#| warning: false
#| message: false

@@

# Render the simplest possible table
knitr::kable(average_table, caption = "Summary Table of Polling Averages by State")

@@

```{r}
#| label: fig-barplot
#| fig-cap: Average Poll Percentage by State
#| echo: false
#| warning: false

@@

logistic_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

@@

### Predicted Election Winner By State
```{r}
#| label: fig-statemap
#| fig-cap: Predicted Election Winner by State
#| echo: false
#| warning: false
#| message: false
#| cache: false

#### Workspace setup ####
library(tidyverse)
library(usmap)
library(ggplot2)
library(rstanarm)

# Check for missing values in state_average and ensure data is clean
state_average <- state_average %>%
  filter(!is.na(Trump_pct) & !is.na(Harris_pct)) %>%
  ungroup()

# Check that state_average has the correct number of rows and no grouping
print(nrow(state_average))  # Ensure this number matches the expected rows

# Proceed with prediction after ensuring the data is clean
predicted_prob <- predict(logistic_model, newdata = state_average, type = "response")

# Ensure predicted_prob matches the number of rows in state_average
if (length(predicted_prob) != nrow(state_average)) {
  stop("Mismatch between predicted_prob and state_average rows")
}

# Add predicted_prob to state_average
state_average <- state_average %>%
  mutate(predicted_prob = predicted_prob)

@@

# Map visualization showing predicted winner by state
plot_usmap(data = state_average, values = "predicted_winner", regions = "states") +
  scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue"), name = "Predicted Winner") +
  labs(title = "Predicted Winner of 2024 US Presidential Election by State") +
  theme(legend.position = "right")

# Remove any states with missing polling data for Trump or Harris
#state_average <- state_average %>%
  #filter(!is.na(Trump_pct) & !is.na(Harris_pct))

# Proceed with the prediction for all states
#state_average <- state_average %>%
  #mutate(predicted_prob = predict(logistic_model, newdata = state_average, type = "response"))

# Classify the predicted winner: Trump if probability > 0.5, else Harris
#state_average <- state_average %>%
  #mutate(predicted_winner = ifelse(predicted_prob > 0.5, "Trump", "Harris"))

#### Map Visualization ####
#plot_usmap(data = state_average, values = "predicted_winner", regions = "states") +
  #scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue"), name = "Predicted Winner") +
  #labs(title = "Predicted Winner of 2024 US Presidential Election by State") +
  #theme(legend.position = "right")
```

@@

knitr::kable(summary_table, caption = "Summary of Predicted 2024 US Presidential Election Results by State", escape = TRUE)

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelsummary
#| tbl-cap: "Explanatory models of election outcome based on polling percentages for Trump and Harris"
#| warning: false

# Ensure the model exists
if (exists("logistic_model")) {
  # Create and display a model summary table in LaTeX format for PDF
  modelsummary::modelsummary(
    list("First Model" = logistic_model),
    statistic = c("std.error", "p.value"),  # Use standard error and p-value
    fmt = 2,  # Format numbers to 2 decimal places
    output = "latex"  # Use LaTeX for PDF output
  )
} else {
  print("Model not found")
}

#### Model Summary Table ####
# Create and display a model summary table
#modelsummary::modelsummary(
  #list("First Model" = logistic_model),
  #statistic = c("std.error", "p.value"),  # Use standard error and p-value
  #fmt = 2,  # Format numbers to 2 decimal places
  #output = "markdown"  # Output format for Quarto or RMarkdown
#)
